---
title: "P8131_hw4_xy2395"
author: "Jack Yan"
date: "2/28/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nnet)
```


## Data Entry
```{r}
data_wide = tibble(
  resp_low = c(65, 34, 130, 141, 67, 130),
  resp_medium = c(54, 47, 76, 116, 48, 105),
  resp_high = c(100, 100, 111, 191, 62, 104),
  contact = rep(c('low', 'high'), 3),
  type = c('tower_block', 'tower_block', 'apartment', 'apartment', 'house', 'house')
)

data_long = 
  data_wide %>% 
  gather(key = 'satisfaction', value = 'count', resp_low:resp_high) %>% 
  mutate(satisfaction = str_sub(satisfaction, 6),
         satisfaction = fct_relevel(satisfaction, 'low', 'medium', 'high'),
         type = fct_relevel(type, 'tower_block','apartment', 'house'),
         contact = fct_relevel(contact, 'low', 'high')
  )

data_wide =  
  data_long %>% 
  spread(satisfaction, count) %>% 
  mutate(
    contact = as_factor(contact),
    type = as_factor(type)
  ) %>% 
  select(low, medium, high, everything())
  
```


## i) Tables

```{r}
tbl1 = 
  data_long %>% 
  group_by(contact, satisfaction) %>% 
  summarize(sum = sum(count)) %>% 
  spread(contact, sum) %>% 
  mutate(high = 100 * high / sum(high),
         low = 100 * low / sum(low)) 

tbl1 %>% knitr::kable(digits = 2,
                      col.names = c('Satisfaction', 'Low Contact (%)', 'High Contact (%)'))
```

The table above shows association between satisfaction and contact with others. Residents with higher degree of contact with others tend to be more satisfied with their housing conditions. The degree of contact is positively associated with housing satisfaction.

```{r}
tbl2 = 
  data_long %>% 
  group_by(type, satisfaction) %>% 
  summarize(sum = sum(count)) %>% 
  spread(type, sum) %>% 
  mutate(tower_block = tower_block / sum(tower_block),
         apartment = apartment / sum(apartment),
         house = house / sum(house)
  )

tbl2 %>% 
  knitr::kable(digits = 2,
               col.names = c('Satisfaction', 'Tower Block (%)', 'Apartment (%)', 'House (%)'))
```

The table above shows association between satisfaction and type of housing. Tower block Residents have the highest proportion of high satisfaction (50%) and the lowest proportion of low satisfaction (25%). Most residents living in apartments have high level of satisfaction, whereas most residents living in houses have low level of satisfaction.



## ii) Multinomial Logistic Regression

```{r}
fit_nomial = multinom(satisfaction ~ contact + type, weights = count, data = data_long)
```

#### Pattern in the associations
```{r}
# Coefficients
summary(fit_nomial)$coefficients
```

The log odds ratio of medium vs low satisfaction between high and low level of contact is 0.296. The log odds ratio of high vs low satisfaction between high and low level of contact is 0.323. This suggests that residents with higher levels of contact have higher level of satisfaction.

The log odds ratio of medium vs low satisfaction between apartment and tower block residents is -0.41. The log odds ratio of high vs low satisfaction between apartment and tower block residents is -0.64. This suggests that people living in tower blocks tends to have the highest level of satisfaction.

#### Odds ratios with 95% confidence intervals
```{r}
fit_nomial %>% 
  broom::tidy() %>% 
  mutate(CIL = estimate * exp(std.error * qnorm(0.025)),
         CIR = estimate / exp(std.error * qnorm(0.025))) %>% 
  select(-std.error, -statistic, -p.value) %>% 
  filter(term != '(Intercept)') %>% 
  knitr::kable(digits = 2)
```


The odds ratios along with their 95% confidence intervals are shown in the table above.

#### Goodness of fit
```{r}
# Goodness of fit
pi_hat = predict(fit_nomial, data_wide, type = 'probs')
# data_long
res = data_wide[,1:3]
m = rowSums(res)
G_stat = sum((res - pi_hat * m)^2 / (pi_hat * m))
G_stat
p_value = 1 - pchisq(G_stat, df = (6-4)*(3-1))
p_value
```

The p-value is `r p_value` > 0.05, so the fit is good.

#### Testing interaction of contact level by house type
```{r}
fit_nomial_interact = multinom(satisfaction ~ contact + type + contact*type, weights = count, data = data_long)
summary(fit_nomial_interact)
D.stat_small = sum(2 * res * log(res/(m*pi_hat)) )
D.stat_small

pi_hat2 = predict(fit_nomial_interact, data_wide, type = 'probs')
D.stat_large = sum(2 * res * log(res/(m*pi_hat2)) )
D.stat_large

# Deviance analysis
pval = 1 - pchisq(D.stat_small - D.stat_large, df = 4)
pval
```

We used deviance analysis to test the interaction between level of contact and housing type. The p-value is larger than 0.05, so we fail to reject the null and conclude that the smaller model (without the interaction) is better. In other words, there is no interaction between contact level and house type.

## iii) Proportional odds model

```{r}
fit_polr = MASS::polr(satisfaction ~ contact + type, data = data_long, weights = count)

summary(fit_polr)
```

The log odds ratio between high and low levels of contact is 0.2524, so people with higher level of contact with others have greater satisfaction. 

The log odds ratio between the housing types apartment and tower block is -0.5009. The log odds ratio between house and tower block is -0.7362. This tells us that house residents have the lowest level of satisfaction and tower block residents have the highest level of satisfaction.

## iv) Pearson residuals

```{r}
pihat_polr = predict(fit_polr, data_wide[,4:5], type = 'prob')

resid_matrix = cbind(data_wide[,4:5], (res - pihat_polr*m)/sqrt(pihat_polr*m))
resid_matrix

max(abs(resid_matrix[,3:5]))
```

The largest Pearson residual is -1.478, corresponding to the response level 'high satisfaction' and covariate levels 'high contact' and 'house'.
