---
title: "p8131_hw6_xy2395"
author: "Jack Yan"
date: "3/31/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nlme)
# dental = read.table('./hw6/HW6-dental.txt')
```

## Problem 2
### 2.1 Spaghetti Plot
```{r}
# Import data
dental <- 
  read.table('HW6-dental.txt', header = TRUE) %>%
  as.tibble() %>% janitor::clean_names() %>% 
  mutate(gender = as.factor(gender))
```

```{r}
# Spaghetti plot
dental %>% 
ggplot(aes(x = age, y = distance, group = child, color = gender)) +
  geom_line() +
  theme_bw()
```

### 2.2 Marginal Form

Answer to this question is hand-written above.

### 2.3 Comparing models with different covariance patterns

For the following 3 models, we assume equal variance across measurements at different ages.

```{r}
# Compound Symmetry covariance
compsym = gls(distance ~ gender + age, 
              data = dental, 
              correlation = corCompSymm(form = ~1 | child), 
              method="REML")
# Exponential covariance
expo = gls(distance ~ gender + age, 
           data = dental, 
           correlation = corExp(form = ~1 | child), 
           method = 'REML')
# Autoregressive covariance
auto1 = gls(distance ~ gender + age, 
            data = dental, 
            correlation = corAR1(form = ~1 | child), 
            method = 'REML')
```

```{r}
# Compare coefficient parameter estimates
bind_rows(
  compsym$coefficients,
  expo$coefficients,
  auto1$coefficients,
) %>% 
  mutate(CovType = c('CompSym', 'Exp', 'Auto')) %>% 
  select(CovType, everything()) %>% 
  knitr::kable()
```

The coefficient parameter estimates are similar across the 3 covariance patterns.

```{r}
# Compare covariance estiamtes
# Compound Symmetry
compsym$sigma^2 * corMatrix(compsym$modelStruct$corStruct)[[1]]
# Exponential covariance
expo$sigma^2 * corMatrix(expo$modelStruct$corStruct)[[1]]
# Autoregressive covariance
auto1$sigma^2 * corMatrix(auto1$modelStruct$corStruct)[[1]]
```

The covariance estiamtes are similar across the 3 covariance patterns.






