---
title: "Untitled"
author: "Jack Yan"
date: "4/15/2019"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nlme)
# polite_df = read_csv('./hw7/HW7-politeness_data.csv') %>% as.tibble() %>% janitor::clean_names()
```

```{r, message = FALSE}
# Import data
polite_df <- 
  read_csv('./hw7/HW7-politeness_data.csv') %>%
  as.tibble() %>% janitor::clean_names() 
```

#### 1. Exploratory Analysis

Provide boxplots to show the relation between gender/attitude and pitch.
```{r}
# boxplots 
polite_df %>%
  ggplot(aes(x = gender, y = frequency, color = attitude)) +
    geom_boxplot() +
    theme_bw()
```

Males generally tend to have lower pitch than females. Within each gender, informal attitude (inf) tends to have higher pitch than formal attitude (pol).

#### 2. Mixed Effects Model with Random Intercept

Fit a mixed effects model with random intercepts for different subjects (gender and attitude being the fixed effects).

```{r}
lmm = lme(frequency ~ gender + attitude, random = ~1 | subject, method = "REML", data = polite_df)

VarCorr(lmm)
# var(Yi) = 598.1953 + 847.7049 = 1445.9
var = as.numeric(VarCorr(lmm)[[1]]) + as.numeric(VarCorr(lmm)[[2]]); var 
# cov(Yij,Yik) = 598.2
cov = as.numeric(VarCorr(lmm)[[1]]); cov 
```

The covariance matrix for a subject $Y_i$ follows a compound symmetry pattern with $var(Y_{ij})=1445.9$ and $cov(Y_{ij},Y_{ik})=598.2$. There are 14 measurements within each subject, so the covariance matrix is a 14*14 matrix with $var(Y_{ij})=1445.9$ as diagonal values and $cov(Y_{ij},Y_{ik})=598.2$ as off-diagonal values.

$var(Y_{i})=\begin{bmatrix} 
    1445.9 &  598.2 & \dots & 598.2 \\
    598.2  & 1445.9 &               \\
    \vdots &        &\ddots & \vdots\\
    598.2  &        & \dots & 1445.9 
 \end{bmatrix}_{14 \times 14}$

```{r}
# covariance matrix for the REML estimates of fixed effects
vcov(lmm)
```


```{r}
# BLUPs for subject-speci􏰍c intercepts
random.effects(lmm)
```