---
title: "p8131_hw10_xy2395"
author: "Jack Yan"
date: "4/30/2019"
output: pdf_document
---
```{r, message=F}
library(tidyverse)
library(survival)
library(survminer)
library(KMsurv)
```


### Problem 1
```{r}
data1 = 
  tibble(
    time = c(4, 12, 15, 21, 23, 2, 6, 8, 10, 19),
    delta = c(1, 0, 1, 0, 1, 1, 0, 0, 1, 1),
    group = c(1, 1, 1, 1, 1, 2, 2, 2, 2, 2)
  )
Surv(data1$time, data1$delta, type = 'right')
# Plot
ggsurvplot(survfit(Surv(time, delta) ~ group, data = data1), conf.int=TRUE)
# Log-rank test
survdiff(Surv(time, delta)~group, data = data1)
```

The chi-squared statistic is 1.1 and p-value is 0.3 >0.05, so we fail to reject the null and conclude that the survival curves of the 2 groups are not different.

### Problem 2
```{r}
data(kidtran)
head(kidtran)
# plot
ggsurvplot(survfit(Surv(time, delta) ~ gender + race, data = kidtran), conf.int=F)
# Log-rank test
# male
survdiff(Surv(time, delta)~ gender + race, data = kidtran, subset = (gender == 1))
# female
survdiff(Surv(time, delta)~ gender + race, data = kidtran, subset = (gender == 2))
```
For males, the chi-squared statistic is 0.1 and p-value is 0.8 >0.05, so we fail to reject the null and conclude that the survival curve of different races is not different in males. On the other hand, for females, the chi-squared statistics is 4.8 and p-value is 0.03 <0.05, so we have enough evidence to reject the null and conclude that the survival curves for different races is different in females.


### Problem 3
```{r}
data(larynx)
head(larynx)
data3 = 
  larynx %>% 
  mutate(z1 = if_else(stage == 2, 1, 0),
         z2 = if_else(stage == 3, 1, 0),
         z3 = if_else(stage == 4, 1, 0)) %>% 
  rename(z4 = age)
fit = coxph(Surv(time, delta)~z1 + z2 + z3 + z4 + z1 * z4, data = data3, ties = "breslow")
summary(fit)
```
The proportional hazards model is:
$$h_i(t)=h_0(t) e^{-7.38Z_1+0.62Z_2+1.75Z_3+0.006Z_4+0.11Z_1 \times Z_4 }$$
Interpretations:

The hazard ratio for subjects of age k in Stage II versus subjects in Stage I of the same age is $e^{-7.38+0.11 \times k}$.

The hazard ratio for subjects in Stage III versus subjects in Stage I is 1.86, keeping age at diagnosis constant.

The hazard ratio for subjects in Stage IV versus subjects in Stage I is 5.77, keeping age at diagnosis constant.

If the patients are not in stage II, the hazard ratio for one-year increase in age at diagnosis is 1.0059908, for patients in the same stage. If the patients are in stage II, the hazard ratio for one-year increase in age at diagnosis is $e^{0.006+0.112}=1.12$.

The relative risk of dying is $e^{-7.38+0.11 \times 50}=0.153$ for a stage II patient of age 50 as compared to a stage I patient of the same age.