---
title: "p8131_hw7_xy2395"
author: "Jack Yan"
date: "4/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nlme)
# polite_df = read_csv('./hw7/HW7-politeness_data.csv') %>% as.tibble() %>% janitor::clean_names()
```

```{r, message = FALSE}
# Import data
polite_df <- 
  read_csv('../hw7/HW7-politeness_data.csv') %>%
  as.tibble() %>% janitor::clean_names() 
```

#### 1. Exploratory Analysis

Provide boxplots to show the relation between gender/attitude and pitch.
```{r}
# boxplots 
polite_df %>%
  ggplot(aes(x = gender, y = frequency, color = attitude)) +
    geom_boxplot() +
    theme_bw()
```

Males generally tend to have lower pitch than females. Within each gender, informal attitude (inf) tends to have higher pitch than formal attitude (pol).

#### 2. Mixed Effects Model with Random Intercept

Fit a mixed effects model with random intercepts for different subjects (gender and attitude being the fixed effects).

```{r}
lmm = lme(frequency ~ gender + attitude, random = ~1 | subject, method = "REML", data = polite_df)

VarCorr(lmm)
# var(Yi) = 598.1953 + 847.7049 = 1445.9
var = as.numeric(VarCorr(lmm)[[1]]) + as.numeric(VarCorr(lmm)[[2]]); var 
# cov(Yij,Yik) = 598.2
cov = as.numeric(VarCorr(lmm)[[1]]); cov 
```

The covariance matrix for a subject $Y_i$ follows a compound symmetry pattern with $var(Y_{ij})=1445.9$ and $cov(Y_{ij},Y_{ik})=598.2$. There are 14 measurements within each subject, so the covariance matrix is a 14*14 matrix with $var(Y_{ij})=1445.9$ as diagonal values and $cov(Y_{ij},Y_{ik})=598.2$ as off-diagonal values.

$var(Y_{i})=\begin{bmatrix} 
    1445.9 &  598.2 & \dots & 598.2 \\
    598.2  & 1445.9 &               \\
    \vdots &        &\ddots & \vdots\\
    598.2  &        & \dots & 1445.9 
 \end{bmatrix}_{14 \times 14}$

```{r}
# covariance matrix for the REML estimates of fixed effects
vcov(lmm)
```

The covariance matrix for the REML estimates of fixed effects is shown above.

```{r}
# BLUPs for subject-speciô°c intercepts
random.effects(lmm)
```

The BLUPs for subject-specific intercepts are shown above. 

```{r}
resid_lmm = polite_df$frequency - fitted(lmm)[[1]]
resid_lmm
```

Residuals are the deviations from subject-specific mean. The residuals are shown above.


#### 3. Likelihood Ratio Test for the Interaction Term

Fit a mixed effects model with intercepts for different subjects (gender, attitude and their interaction being the fixed effects).

```{r}
lmm_small = lme(frequency ~ gender + attitude, random = ~1 | subject, method = "ML", data = polite_df)
lmm_large = lme(frequency ~ gender * attitude, random = ~1 | subject, method = "ML", data = polite_df)

anova(lmm_small, lmm_large)
```

We use maximum likelihood method to fit the model. The p-value for Likelihood ratio test is 0.2392 > 0.05, so we fail to reject the null and conclude that the interaction term is not significantly associated with pitch, at the significance level of 0.05.


#### 4. Mixed Effects Model with Random Intercept and Slope

Model: $Y_{ij}=\beta_1 + \beta_2*I(gender=Female)+\beta_3*t_{ij}+b_{1i}+b_{2i}*t_{ij}+ \epsilon_{ij}$, where $t_{ij}=I(attitude_{ij}=pol)$, $\beta_1$ is the fixed intercept, $\beta_2$ and $\beta_3$ are slopes, and $b_{1i}$ and $b_{2i}$ are random intercept and random slope, respectively.

```{r}
lmm2 = lme(frequency ~ gender + attitude, random = ~1 + attitude | subject, method = "REML", data = polite_df)

VarCorr(lmm2)
getVarCov(lmm2)
```
We get $g_{11}=598.2$, $g_{22}=1.079496*10^{-5}$, $g_{12}=0$, and $\sigma^2=847.7$. 

Variance:
$var(Y_{ij})=g_{11} + 2t_{ij}g_{12} + t_{ij}^2 g_{22} + \sigma^2$

1. If $t_{ij}= 1$, $var(Y_{ij})=g_{11} + 2t_{ij}g_{12} + t_{ij}^2 g_{22} + \sigma^2=598.2+2*0+1.079496*10^{-5}+847.7=1445.9$

2. If $t_{ij}= 0$, $var(Y_{ij})=g_{11} + \sigma^2=598.2+847.7=1445.9$

The general formula of covariance between $Y_{ij}$ and $Y_{ik}$ is

$cov(Y_{ij},Y_{ik})=g_{11}+ (t_{ij} + t_{ik})g_{12} +t_{ij}t_{ik}g_{22}$.

1. If $t_{ij}= 0$ and $t_{ik}=0$, $cov(Y_{ij},Y_{ik})=g_{11}= 598.2$.

2. If $t_{ij}= 0$ and $t_{ik}=1$, $cov(Y_{ij},Y_{ik})=g_{11}+g_{12}=598.2+0=598.2$.

3. If $t_{ij}= 1$ and $t_{ik}=0$, $cov(Y_{ij},Y_{ik})=g_{11}+g_{12}=598.2+1.079496*10^{-5}=598.2$.

4. If $t_{ij}= 1$ and $t_{ik}=1$, $cov(Y_{ij},Y_{ik})=g_{11}+ 2g_{12}+g_{22}=598.2+2*0+1.079496*10^{-5}=598.2$.

So we get $var(Y_{ij})= 1445.9$ and $cov(Y_{ij},Y_{ik})=598.2$ for any different j and k.
**Matrix Notation:**

Suppose there are 14 observations within an individual, and for observations 1-7 $t_{ij}=0$, and for observations 8-14 $t_{ij}=1$. The covariance matrix is as follows:

$A=\begin{bmatrix}
g_{11} + \sigma^2 & g_{11} & \dots & g_{11} \\
g_{11} & g_{11} + \sigma^2 &  & \vdots \\
\vdots &  & \ddots & g_{11} \\
g_{11} & \dots & g_{11} & g_{11} + \sigma^2 
\end{bmatrix}_{7 \times 7}$,

$B=\begin{bmatrix}
g_{11} + 2t_{ij}g_{12} + t_{ij}^2 g_{22} + \sigma^2 & g_{11}+ 2g_{12}+g_{22} & \dots & g_{11}+ 2g_{12}+g_{22} \\
g_{11}+ 2g_{12}+g_{22} & g_{11} + 2t_{ij}g_{12} + t_{ij}^2 g_{22} + \sigma^2 &  & \vdots \\
\vdots &  & \ddots & g_{11}+ 2g_{12}+g_{22} \\
g_{11}+ 2g_{12}+g_{22} & \dots & g_{11}+ 2g_{12}+g_{22} & g_{11} + 2t_{ij}g_{12} + t_{ij}^2 g_{22} + \sigma^2 
\end{bmatrix}_{7 \times 7}$,

$C=\begin{bmatrix}
g_{11}+g_{12} & g_{11}+g_{12} & \dots & g_{11}+g_{12} \\
g_{11}+g_{12} & g_{11}+g_{12} &  & \vdots \\
\vdots &  & \ddots & g_{11}+g_{12} \\
g_{11}+g_{12} & \dots & g_{11}+g_{12} & g_{11}+g_{12} 
\end{bmatrix}_{7 \times 7}$

$var(Y_{i})= \begin{bmatrix}
A & C \\
C & B 
\end{bmatrix}_{14 \times 14}=\begin{bmatrix} 
    1445.9 &  598.2 & \dots & 598.2 \\
    598.2  & 1445.9 &               \\
    \vdots &        &\ddots & \vdots\\
    598.2  &        & \dots & 1445.9 
 \end{bmatrix}_{14 \times 14}$

The random slope has small variance, so it captures little variation in the data. So this model does not do a much better job than a random intercept model. The covariance structure for this model is thus numerically a compound symmetry structure as in Question 2.

```{r}
fixed.effects(lmm2)
fixed_effect = 256.98691 + -20.00238; fixed_effect

random.effects(lmm2)
random_effect = -13.575831 + -8.408891 * 10^(-07); random_effect

BLUP = fixed_effect + random_effect; BLUP
```
The fixed effect for intercept is 256.98691, and for attitude being polite is -20.00238; so the fixed effects in total is 236.9845. The random effect for intercept is -13.575831 and for attitude being polite is $-8.408891 \times 10^{-07}$; so the random effects in total is -13.57583. The BLUP is the summation of fixed effects and random effects, 223.4087.

#### 5. ?
```{r}
library(lme4)
lmm3 = lmer(frequency ~ gender + attitude + (1 | subject) +  (1 | scenario), data = polite_df)
summary(lmm3)
# Variance of Y_ij
224.5 + 613.2 + 637.8
# Covariance of Y_ij and Y_ik with same scenario
224.5 + 613.2
# Covariance of Y_ij and Y_ik with different scenarios
613.2
```

Denote variance of subject specific intercept as $\sigma_1^2$, variance of scenario specific intercept as $\sigma_2^2$, and variance of residuals as $\sigma^2$. We have $\sigma_1^2=613.2$, $\sigma_2^2=224.5$, and $\sigma^2=637.8$.

$A=
\begin{bmatrix}
\sigma_1^2+\sigma_2^2+\sigma^2 & \sigma_1^2+\sigma_2^2 \\
\sigma_1^2+\sigma_2^2 & \sigma_1^2+\sigma_2^2+\sigma^2 
\end{bmatrix}=
\begin{bmatrix}
1475.5 & 837.7 \\
837.7 & 1475.5 
\end{bmatrix}$ is the variance-covariance matrix of 2 observations with the same scenario within an individual. 

$B=\begin{bmatrix}
\sigma_1^2 & \sigma_1^2 \\
\sigma_1^2 & \sigma_1^2 
\end{bmatrix}=
\begin{bmatrix}
613.2 & 613.2 \\
613.2 & 613.2 
\end{bmatrix}$ is the covariance matrix of observations with different scenarios within an individual.

We get $var(Y_i)=\begin{bmatrix}
A & B & \dots & B \\
B & A &  & \vdots \\
\vdots &  & \ddots & B \\
B & \dots & B & A
\end{bmatrix}_{14 \times 14}$.

Interpretation:

The mean pitch will on avarage be 20.002 units lower for polite attitude than informal attitude over all observations, within the same gender.




