---
title: "p8131_hw9_xy2395"
author: "Jack Yan"
date: "4/23/2019"
output: pdf_document
---


### Problem 1

We have $$h(t)=\frac{2t}{1+t^2}=\frac{d}{dt}\lbrack log(1+t^2) \rbrack=\frac{d}{dt} \lbrack -logS(t) \rbrack$$
$$log(1+t^2) = -logS(t)$$
Therefore, the survival function is $$S(t) = \frac{1}{1+t^2}$$
The cumulative density function is $$F(t) = 1 - S(t) = 1 - \frac{1}{1+t^2}$$
The probability density function is $$f(t) = \frac{d}{dt}F(t) = \frac{2t}{(1+t^2)^2}$$

### Problem 2
```{r, include = F}
library(tidyverse)
tibble(
  t_i = c(1,2,4,5,6,7,8,9,10),
  n_i = c(10,9,7,6,5,4,3,2,1),
  d_i  = c(1,2,0,0,1,0,0,0,0),
  c_i = c(0,0,1,1,0,1,1,1,1)
) %>%
knitr::kable()
```

|$t_i$|$n_i$|$d_i$|$c_i$|$\hat{\lambda_i}$|                                $\hat{S}(t)$|                           $\tilde{H}(t)$| $exp(-\tilde{H}(t))$|
|----:|----:|----:|----:|----------------:|-------------------------------------------:|----------------------------------------:|--------------------:|
|    1|   10|    1|    0|   $\frac{1}{10}$|           $1*(1-\frac{1}{10})=\frac{9}{10}$|                           $\frac{1}{10}$|0.905|
|    2|    9|    2|    0|    $\frac{2}{9}$| $\frac{9}{10}*(1-\frac{2}{9})=\frac{7}{10}$| $\frac{1}{10}+\frac{2}{9}=\frac{29}{90}$|0.725|
|    4|    7|    0|    1|                0|                              $\frac{7}{10}$|                          $\frac{29}{90}$|0.725|
|    5|    6|    0|    1|                0|                              $\frac{7}{10}$|                          $\frac{29}{90}$|0.725|
|    6|    5|    1|    0|    $\frac{1}{5}$|$\frac{7}{10}*(1-\frac{1}{5})=\frac{14}{25}$|$\frac{29}{90}+\frac{1}{5}=\frac{47}{90}$|0.593|
|    7|    4|    0|    1|                0|                             $\frac{14}{25}$|                          $\frac{47}{90}$|0.593|
|    8|    3|    0|    1|                0|                             $\frac{14}{25}$|                          $\frac{47}{90}$|0.593|
|    9|    2|    0|    1|                0|                             $\frac{14}{25}$|                          $\frac{47}{90}$|0.593|
|   10|    1|    0|    1|                0|                             $\frac{14}{25}$|                          $\frac{47}{90}$|0.593|


### Problem 3
```{r setup, include=TRUE, message=F, warning=F}
library(tidyverse)
library(KMsurv) # data source
library(survival) # problem 3
```

```{r}
data(tongue)
tongue_df = 
  tongue %>% as.tibble() %>% 
  mutate(type = if_else(type == 1, "Aneuploid", "Diploid"))
```

```{r}
# Aneuploid K-M curve
KM1 = survfit(Surv(time,delta)~1, data = subset(tongue_df, type == "Aneuploid"), conf.type = 'log')
plot(KM1, 
     conf.int = TRUE, mark.time = TRUE, 
     xlab = "Weeks", ylab = "Survival Probability", main="Aneuploid K-M curve", 
     cex.lab = 1.5, cex.main = 1.5)

# Diploid K-M curve
KM2 = survfit(Surv(time,delta)~1, data = subset(tongue_df, type == "Diploid"), conf.type = 'log')
plot(KM2, 
     conf.int = TRUE, mark.time = TRUE, 
     xlab = "Weeks", ylab = "Survival Probability", main="Diploid K-M curve", 
     cex.lab = 1.5, cex.main = 1.5)
``` 

```{r}
# one-year survival rate
# Aneuploid
summary(KM1, time = 365/7)
# Diploid
summary(KM2, time = 365/7)
```

The estimated 1-year survival rate (95% CI) is 0.654 (0.537, 0.797) for Aneuploid and 0.486 (0.33, 0.716) for Diploid. 
