---
title: "P8131_hw5_xy2395"
author: "Jack Yan"
date: "3/5/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

```{r echo = T, eval = F}
crab_df = 
  read.table("./hw5/HW5-crab.txt", header = TRUE) %>% 
  as.tibble()
```

```{r echo = F, eval = T}
crab_df = 
  read.table("./HW5-crab.txt", header = TRUE) %>% 
  as.tibble()
```

### (a)

M1 : Sa ~ W

```{r}
fit1 = glm(Sa ~ W, family = poisson, data = crab_df)

fit1 %>% 
  broom::tidy() %>% 
  mutate(exp_estimate = exp(estimate)) %>% 
  select(-std.error, -statistic, -p.value) %>% 
  knitr::kable()

# Deviance and Pearson statistics are similar
D = sum(residuals(fit1, type = 'deviance')^2); D
G = sum(residuals(fit1, type = 'pearson')^2); G

1 - pchisq(G, df = fit1$df.residual) # df = 171
```

The p-value is lower than 0.05, and the deviance from the full model is too large, so the fit is not good.

**Interpretation**: The estimated coefficient of W (carapace width) is 0.164. exp(0.164) = 1.18. This means the number of satellites is estimated to increase by 1.18 folds with one-unit increase in carapace width.

```{r eval = F, echo = F}
D = sum(residuals(fit1, type = 'deviance')^2)
D = fit1$null.deviance - fit1$deviance # This is not correct!!!
G = sum(residuals(fit1, type = 'pearson')^2)
G

# alternative way to calculate pearson statistics
resp = predict.glm(fit1, type = 'response')
G = sum((crab_df$Sa - resp)^2 / resp )
```


### (b)

M2: Sa ~ W + Wt

```{r}
fit2 = glm(Sa ~ W + Wt, family = poisson, data = crab_df)

fit2 %>% 
  broom::tidy() %>% 
  mutate(exp_estimate = exp(estimate)) %>% 
  select(term, estimate, exp_estimate, p.value) %>% 
  knitr::kable()

test.stat = fit1$deviance - fit2$deviance
df = fit1$df.residual - fit2$df.residual
1 - pchisq(test.stat, df = df) # df = 1
```

The p-value is smaller than 0.05, so we reject the null hypothesis and conclude that the larger model (M2) is superior to the smaller model (M1). 

### (c)

Check Over Dispersion.
```{r, warning=F}
resid = residuals(fit2, type = 'pearson')
phi = sum(resid^2) / fit2$df.residual
phi

n = fit2$df.null + 1
plot(qnorm((n + 1:n + 0.5)/(2*n + 1.125)), sort(abs(resid)),
     xlab = 'Expected Half-Normal Order Stats',
     ylab = 'Ordered Abs Pearson Residuals')
abline(a = 0, b = 1)
abline(a = 0, b = sqrt(phi), lty = 2)

summary(fit2, dispersion = phi)$coef %>%
  broom::tidy() %>% 
  mutate(exp_estimate = exp(Estimate)) %>% 
  select(.rownames, Estimate, exp_estimate) %>%
  knitr::kable()
```

As shown in the plot, there is over-dispersion in Model 2. 

**Interpretation**: The number of satellites is estimated to increase by 1.05 folds with one-unit increase in female crab's carapace width. The number of satellites is estimated to increase by 1.56 folds with one-unit increase in female crab's weight.



## Problem 2

```{r}
parasite_df = 
  read.table("./HW5-parasite.txt", header = TRUE) %>% 
  as.tibble()
```

